Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire

### 2016-04-11 manuscript/supplementary.Rnw
Fixed formatting of figures and other small changes

\subsection*{Data inputs for growth, mortality and recruitment models.}
In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some distances were censored (i.e. $\geq$ 26 cm), these data were incomplete. We estimated these censored data using a censored model (see Modelling censored Poa Distances). Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.

@@

\subsection{\textit{Data inputs into density and maximum height models.}}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity ($Severity_j$), was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were \textit{r} \textless 0.16. The data from each continuous covariate was then centered on their mean and standardized by two standard deviations\cite{Gelman:2007te}.

\section*{Data analysis}
In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}\cite{stan:2015xx}. For each model, three independent chains were run. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}. In all cases modelled parameters converged within 1000 iterations. After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

\subsection{\textit{Modelling shrub seedling growth rates.}}
We modelled the observed height of individual $i$ at May census $t$ as a random realisation from a normal distribution truncated at zero:

@@

\begin{equation}\label{eq:r_tussock} 
\begin{aligned}
R_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
\end{aligned}
\end{equation}

@@

\subsection{\textit{Modelling shrub seedling mortality rates.}}
We were also interested in whether the instantaneous mortality rates of seedlings varied amongst shrub species, OTC treatment and inter-tussock gap size. Our observations, $y_{i,t}$, were binary (0 = alive or 1 = dead) for each individual $i$, in each census, $t$. We modelled these binary observations through time as random realisations from a Bernoulli distribution based on the model-estimated cumulative hazard rate for each individual at each census:

@@

\begin{equation}\label{eq:mort_tussock} 
\begin{aligned}
\lambda_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
\end{aligned}
\end{equation}

@@

\subsection{\textit{Modelling Grevillea recruitment rates.}}
We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realisation from a negative binomial distribution:

@@

\subsection{\textit{Modelling censored Poa distances.}}
Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ distance data used to calculate the median inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e. $\geq$  26 cm).  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects, and then using these distributions to estimate censored observations.

@@

\begin{equation}\label{eq:censored_obs} 
censored\;distance_i \sim LN(\mu_i, \sigma_{obs}), 26 \leq censored\;distance_i \leq 100.
\end{equation}

@@

\subsection{\textit{Modelling rates of inter-tussock gap infilling by Poa.}}
As well as modelling how OTCs affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps. To examine this we first modelled the median observed gap radius, $y_{i,t}$, for inter-tussock gap $i$ in census $t$, as a random realisation from a lognormal distribution:

@@

\subsection{\textit{Modelling drivers of seedling density.}}
We examined drivers of seedling density by modelling the number of seedlings per square metre plot $i$ as a random realisation from a Poisson distribution:

@@

\begin{equation}\label{eq:density_site} 
\begin{aligned}
\delta_j &= \alpha_j + \beta_1 \times Alt_j + \beta_2 \times TWI_j + \beta_3 \times (1 - Burnt_j) + \beta_4 \times (Severity_j \times Burnt_j) \\
\alpha_j &\sim \mathcal{N}(\mu_{site}, \sigma_{site}),
\end{aligned}
\end{equation}

@@

\subsection*{\textit{Modelling drivers of seedling maximum heights.}}
In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realisations from a log-normal distribution:

>>>>

\subsubsection*{Data inputs for growth, mortality and recruitment models}
In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some distances were censored (i.e. $\geq$ 26 cm), these data were incomplete. We estimated these censored data using a censored model (see Modelling censored Poa Distances). Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.

@@

\subsubsection*{Data inputs into density and maximum height models}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity ($Severity_j$), was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were \textit{r} \textless 0.16. The data from each continuous covariate was then centered on their mean and standardized by two standard deviations\cite{Gelman:2007te}.

\subsection*{Data analysis}
In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}\cite{stan:2015xx}. For each model, three independent chains were run. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}. In all cases modelled parameters converged within 1000 iterations. After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

\subsubsection*{Modelling shrub seedling growth rates}
We modelled the observed height of individual $i$ at May census $t$ as a random realisation from a normal distribution truncated at zero:

@@

\begin{equation}\label{eq:r_tussock} 
\begin{aligned}
R_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} \\
& + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
\end{aligned}
\end{equation}

@@

\subsubsection*{Modelling shrub seedling mortality rates}
We were also interested in whether the instantaneous mortality rates of seedlings varied amongst shrub species, OTC treatment and inter-tussock gap size. Our observations, $y_{i,t}$, were binary (0 = alive or 1 = dead) for each individual $i$, in each census, $t$. We modelled these binary observations through time as random realisations from a Bernoulli distribution based on the model-estimated cumulative hazard rate for each individual at each census:

@@

\begin{equation}\label{eq:mort_tussock} 
\begin{aligned}
\lambda_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} \\
& + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
\end{aligned}
\end{equation}

@@

\subsubsection*{Modelling Grevillea recruitment rates}
We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realisation from a negative binomial distribution:

@@

\subsubsection*{Modelling censored Poa distances}
Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ distance data used to calculate the median inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e. $\geq$  26 cm).  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects, and then using these distributions to estimate censored observations.

@@

\begin{equation}\label{eq:censored_obs} 
censored_i \sim LN(\mu_i, \sigma_{obs}), 26 \leq censored_i \leq 100.
\end{equation}

@@

\subsubsection*{Modelling rates of inter-tussock gap infilling by Poa}
As well as modelling how OTCs affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps. To examine this we first modelled the median observed gap radius, $y_{i,t}$, for inter-tussock gap $i$ in census $t$, as a random realisation from a lognormal distribution:

@@

\subsubsection*{Modelling drivers of seedling density}
We examined drivers of seedling density by modelling the number of seedlings per square metre plot $i$ as a random realisation from a Poisson distribution:

@@

\begin{equation}\label{eq:density_site} 
\begin{aligned}
\delta_j &= \alpha_j + \beta_1 \times Alt_j + \beta_2 \times TWI_j + \beta_3 \times (1 - Burnt_j) \\
& + \beta_4 \times (Severity_j \times Burnt_j) \\
\alpha_j &\sim \mathcal{N}(\mu_{site}, \sigma_{site}),
\end{aligned}
\end{equation}

@@

\subsubsection*{Modelling drivers of seedling maximum heights}
In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realisations from a log-normal distribution:
